---
title: "Problem Set 1"
subtitle: "Daniel Shapiro"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(6800)
library(tidyverse)
library(stats)
```

### 1a: What is the overall probability of being born in August?

```{r 1a}
# First, let's define what we already know.
P( )
```

### 1b: What is P(Name=$Gus$|Man born in August)? That is, conditional on being born in August, what is the probability that a man is named Gus?

### 1c: What is P(Name=$Gus$|Man born in July)?

### 1d: Imagine that there are 600 men born in 1982. Given the conditions above, what is the probability that at least two baby $Guses$ are born on August 1, 1982? 

### 1e: What is the probability that at least two baby $Hugos$ are born on August 1, 1982?

### 1f: Given your answers to (d) and (e), what would you say to a pundit who pointed to two $Guses$ with the same August 1, 1982 birthday as evidence of voter fraud in the form of double voting? What would you say to a pundit who pointed to two $Hugos$ with the same August 1, 1982 birthday as evidence of voter fraud in the form of double voting?